Spin canonical divisors of log stacky curves

### 2015-06-29 stacky_write_up/spin_cring.tex
added extra end defn
section 3 edits???

\begin{cor}
\label{cor:main-mod-forms}
Let $\Gamma$ be a Fuchsian group and $\sx$ the stacky curve
associated to $\Gamma \backslash \BH$ with signature $(g; e_1,
\ldots, e_r; \delta)$. Then the ring of modular forms
$M(\Gamma) := \bigoplus_{k = 0}^{\infty} M_k(\Gamma)$ is
generated as a $\BC$-algebra by elements of degree at most $e = \max
 (5$, $e_1$, $\ldots$, $e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$
then $M(\Gamma)$ has relations generated in degrees up to $\max(11,
2e_1, \ldots, 2e_r)$. Otherwise, $M(\Gamma)$ has relations
generated in degrees up to $\max(10, 2e_1, \ldots, 2e_r)$, so long
as $\sigma$ does not lie in a finite list of exceptional cases, as
listed in Table ~\ref{table:g-1-exceptional} for signatures with $g
= 1$ and Table ~\ref{table:g-0-exceptional} for signatures with $g =
0$.
\end{cor}

@@

\section{Background}
\label{sec:background}
Here we denote various definitions and useful notation that will be
used throughout the paper. Many of the conventions come from existing
literature. For basic references on the statements and definitions
used below, see Hartshorne \cite[Chapter IV]{hartshorne:ag},
Saint-Donat \cite{saint-donat:proj}, the book of Arbarello--
Cornalba--Griffiths--Harris \cite[Section III.2]
{acgh:algebraic-curves}, and Voight--Zureick-Brown \cite[Chapter 2,
Chapter 5]{vzb:stacky}.

\ssec{Stacky Curves and Log Spin Canonical Rings}
\label{ssec:stacky-background}
A \textbf{stacky curve} $\sx$ over a field $k$ is a smooth proper
geometrically-connected Deligne-Mumford stack of dimension $1$ over
$k$ with a dense open subscheme. A stacky curve is \textbf{tame} if
its stabilizers are coprime to char $k$. Throughout this paper,
stacky curves will be assumed to be tame and $k$ will always be
algebraically closed. The language of stacks will be omitted in
this paper, as it is sufficient to view stacky curves and related
objects $\sx$ as a scheme whose associated divisor group allows for
a finite number of \textbf{stacky points} which may be viewed
heuristically as ``fractional points''. See Voight and
Zureick-Brown for more thorough treatment of these definitions
\cite[Chapter 5]{vzb:stacky}.

@@

\todo{Note that this result is independence of base field}

If $R$ is a graded ring, then $(R)_d$ refers to the $d^{th}$ degree
component of $R$.

@@

\ssec{Monomial Ordering}
\label{ssec:monomial-order}
Here we give a brief overview of the three monomial orderings
that we use. For further reference on monomial orderings,
initial ideals, and Gr\"{o}bner bases, see Eisenbud
\cite[Section 15.9]{eisenbud:comm-alg} and Cox--Little--Shea
\cite[Chapter 2]{cls:ideals-varieties-algorithms}.

\begin{defn}
We equip the polynomial ring $k[x_1, \ldots, x_n]$ with the
(weighted graded) reverse lexicographic order \textbf{grevlex}
$\prec$: if 
\[
	x^{\vec{m}} = \prod_{d,i} x_{d,i}^{m_{d,i}}
\]

\noindent
and $x^{\vec{n}}$ are monomials in $k[x]_{\vec{a}}$, then $x^{\vec{m
}} \succ x^{\vec{n}}$ if and only if either
\begin{equation}
\vec{a}\cdot\vec{m} = \sum_{d,i} dm_{d,i}  > \vec{a} \cdot \vec{n}
\end{equation}
or
\begin{equation} \label{eqn:grevlexdef}
\text{$\vec{a}\cdot\vec{m} = \vec{a} \cdot \vec{n}$ and the last (right-most) nonzero entry in $\vec{m}-\vec{n}$ is negative.}
\end{equation}
It is important to note that in \eqref{eqn:grevlexdef}, the ordering of the variables matters: it corresponds to a choice of writing the exponents of a monomial as a vector.  A common choice for us will be
\begin{equation} \label{eqn:xdOrder}
 x_{1,1}^{m_{1,1}} \cdots x_{D,a_D}^{m_{D,a_D}} \leftrightarrow (m_{D,1},\dots,m_{D,a_D},\ldots,m_{1,1},\dots,m_{1,a_1}) 
 \end{equation}
in which case we have $x_{2,1} \succ x_{1,2}^2$
and $x_{5,1} \succ x_{1,6}^2 x_{3,2}$, etc.
We indicate this ordering in the presentation of the ring, e.g.~for the above we would write
\[ k[x]_{\vec{a}}= k[x_{1,1},\dots,x_{1,a_1},\dots,x_{D,1},\dots,x_{D,a_D}]. \]
In this way, our relations write generators in larger degree in terms of those in smaller degree (``later generators in terms of earlier generators''), which gives the most natural-looking canonical rings to our eyes.

For a homogeneous polynomial $f \in k[x]_{\vec{a}}$, we define the \textbf{initial term} $\initial_{\prec}(f)$ to be the monomial in the support of $f$ that is largest with respect to $\prec$, and we define $\initial_{\prec}(0)=0$.  Let $I \subseteq k[x]_{\vec{a}}$ be a homogeneous ideal.  We define the \textbf{initial ideal} $\initial_{\prec}(I)$ to be the ideal generated by $\{\initial_{\prec}(f) : f \in I\}$.  A \textbf{Gr\"obner basis} (also called a \textbf{standard basis}) for $I$ is a set $\{f_1,\dots,f_n\} \subset I$ such that 
\[ \langle \initial_{\prec}(f_1),\dots,\initial_{\prec}(f_n) \rangle = \initial_{\prec}(I). \]

\begin{rem}
A Gr\"obner basis for $I$ \emph{a priori} depends on a choice of basis for the ambient graded polynomial ring.
\end{rem}

We also introduce a term ordering that is well-suited for inductive arguments: the \emph{block} term order.  
In our inductive arguments, we will often have the following setup: an inclusion $R \supset R'$ of canonical rings such that $R'$ is generated by elements $x_{i,d}$ and $R$ is generated over $R$ by elements $y_j$.  It is natural, therefore, to consider term orders which treat these sets of variables separately.  More formally, we make the following definition.

Let $k[x]_{\vec{a}}$ and $k[y]_{\vec{b}}$ be weighted polynomial rings with term orders $\prec_x$ and $\prec_y$, respectively, and consider $k[y,x]_{\vec{b},\vec{a}} = k[y]_{\vec{b}} \otimes_k k[x]_{\vec{a}}$ the common weighted polynomial ring in these two sets of variables $y,x$.

\begin{defn}
The \textbf{(graded) block} (or \textbf{elimination}) term ordering on $k[y,x]_{\vec{b},\vec{a}}$ is defined as follows: we declare
  \[ y^{\vec{m}}x^{\vec{n}}  \succ y^{\vec{m}'}x^{\vec{n}'}\]  
  if and only if
  \begin{enumerate}
  \item[(i)] $\deg y^{\vec{m}} x^{\vec{n}}  > \deg y^{\vec{m}'} x^{\vec{n}'} $, or
  \item[(ii)] $\deg y^{\vec{m}}x^{\vec{n}} = \deg y^{\vec{m}'} x^{\vec{n}'} $ and 
    \begin{enumerate}
    \item $y^{\vec{m}}\succ_y y^{\vec{m}'}$ or
    \item $y^{\vec{m}} =  y^{\vec{m}'}$ and $x^{\vec{n}} \succ_x x^{\vec{n}'}$.
    \end{enumerate}
  \end{enumerate}
\end{defn}

The block ordering is indeed a term order: the displayed inequalities directly give that any two monomials are comparable, and the inequalities are visibly stable under multiplication by a monomial.  One can similarly define an \textbf{iterated (graded) block} term ordering for any finite number of weighted polynomial rings.

The block ordering is the most suitable ordering for the structure of $R$ as an $R'$-algebra

background on Gr\"{o}bner bases
Notation for Gr\"{o}bner bases and orderings (include grlex, grevlex, and block order, as defined in \cite[Definition 8.1.1]{vzb:stacky})

@@

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$ in degrees at most $\beta$ with $I$ generated over $I'$ in degrees at most $\max(2\beta, \beta + \tau)$ (where $\tau = \max(1, \max_{1\le i, m}(\deg(x_i)))$.  \todo{Peter: Part (c) is immediate from (a) and (b), so maybe it shouldn't be its own section.  On the other hand, the other parts already have a lot of information written down (especial part (b)), so its not clear where it should be put.  Maybe in a remark after the lemma?}
%\item If $\gin_\prec(I')=\initial_\prec(I')$ then $\gin_\prec(I)=\initial_\prec(I)$.
%\item[(c)] A Gr\"{o}bner basis with the initial terms $y_i x_j$, $y_i y_j$,
%	and $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
%	\frac{\alpha}{\beta} \rfloor} + 1}$ (really the subset of these with positive subscripts) as shown above minimally
%generates $I'$.
\end{enumerate}
\end{lem}
\begin{proof}
By
Equation ~\ref{eqn:deg1-sat-ind}, for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

\ssec{Raising Stabilizer Orders}
\label{ssec:raise-orders}
In this subsection, we aim to prove Lemma
~\ref{lem:raise-stacky-order}, whose proof is almost identical to that of \cite[Theorem 8.5.7]{vzb:stacky}. In many cases, Lemma ~\ref{lem:raise-stacky-order} implies that if the main result, Theorem 
~\ref{thm:main}, holds for a curve with signature
$(g;e_1', \ldots, e_k', e_{k+1}' \ldots, e_r';\delta),$ with $e_{k+1}' = \cdots = e_r',$ then Theorem ~\ref{thm:main} also holds for a curve with
signature $(g; e_1', \ldots, e_k',e_{k+1}'+2,\ldots, e_r'+2; \delta).$ In order to state Lemma
~\ref{lem:raise-stacky-order}, we will need to define a certain notion of admissibility in Definition ~\ref{defn:admissible}, which is quite similar to \cite[Definition 8.5.1]{vzb:stacky}.

One key difference between the notion of admissibility in Definition ~\ref{defn:admissible} and \cite[Definition 8.5.1]{vzb:stacky} is that in \cite{vzb:stacky}, Voight and Zureick-Brown may assume that $\{P_i\} \cap \Supp(K_X) = \emptyset.$ However, we cannot assume that $\{P_i\} \cap \Supp(L_X) = \emptyset$, as $L_X$ is not necessarily basepoint-free. Therefore, we will have to work with orders of zeros of poles relative to the divisor we are inducting from, instead of relative to the structure sheaf.

>>>>

\begin{cor}
\label{cor:main-mod-forms}
Let $\Gamma$ be a Fuchsian group and $\sx$ the stacky curve
associated to $\Gamma \backslash \BH$ with signature $(g; e_1,
\ldots, e_r; \delta)$. Then the ring of modular forms
$M(\Gamma) := \bigoplus_{k = 0}^{\infty} M_k(\Gamma)$ is
generated as a $\BC$-algebra by elements of degree at most $e = \max
 (5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$
then $M(\Gamma)$ has relations generated in degrees up to $\max(11,
2e_1, \ldots, 2e_r)$. Otherwise, $M(\Gamma)$ has relations
generated in degrees up to $\max(10, 2e_1, \ldots, 2e_r)$, so long
as $\sigma$ does not lie in a finite list of exceptional cases, as
listed in Table ~\ref{table:g-1-exceptional} for signatures with $g
= 1$ and Table ~\ref{table:g-0-exceptional} for signatures with $g =
0$.
\end{cor}

@@

\section{Background}
\label{sec:background}
Here we denote various definitions and useful notation that will be
used throughout the paper. Many of the conventions come from existing
literature. In particular, we use much of the notation from Voight
and Zureick-Brown. Here, we briefly present the definitions
and notation we use.

A \textbf{stacky curve} $\sx$ over a field $k$ is a smooth proper
geometrically-connected Deligne-Mumford stack of dimension $1$ over
$k$ with a dense open subscheme. A stacky curve is \textbf{tame} if
its stabilizers are coprime to char $k$. Throughout this paper,
stacky curves will be assumed to be tame and $k$ will always be
algebraically closed. The language of stacks will be omitted in
this paper, as it is sufficient to view stacky curves and related
objects $\sx$ as a scheme whose associated divisor group allows for
a finite number of \textbf{stacky points} which may be viewed
heuristically as ``fractional points''. See Voight and
Zureick-Brown for more thorough treatment of these definitions
\cite[Chapter 5]{vzb:stacky}.

@@

\begin{enumerate}
	\item Note that this result is independence of base field
	\item background on Gr\"{o}bner bases
	\item Notation for Gr\"{o}bner bases and orderings (include grlex, grevlex, and block order, as defined in \cite[Definition 8.1.1]{vzb:stacky})
\end{enumerate}

Define $(R)_d$ to refer to the $d^{th}$ component

@@

@@

\item[(c)] In particular, $R_\halfcan$ is generated over $R_\halfcan'$ in degrees at most $\beta$ with $I$ generated over $I'$ in degrees at most $\max(2\beta, \beta + \tau)$ (where $\tau = \max(1, \max_{1\le i, m}(\deg(x_i)))$).  \todo{Peter: Part (c) is immediate from (a) and (b), so maybe it shouldn't be its own section.  On the other hand, the other parts already have a lot of information written down (especial part (b)), so its not clear where it should be put.  Maybe in a remark after the lemma?}
%\item If $\gin_\prec(I')=\initial_\prec(I')$ then $\gin_\prec(I)=\initial_\prec(I)$.
%\item[(c)] A Gr\"{o}bner basis with the initial terms $y_i x_j$, $y_i y_j$,
%	and $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor
%	\frac{\alpha}{\beta} \rfloor} + 1}$ (really the subset of these with positive subscripts) as shown above minimally
%generates $I'$.
\end{enumerate}
\end{lem}
\begin{proof}
By
Equation ~\ref{eqn:deg1-sat-ind}, for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

\ssec{Raising Stabilizer Orders}
\label{ssec:raise-orders}
In this subsection, we present Lemma
~\ref{lem:raise-stacky-order}, whose proof is almost identical to one of Voight and Zureick-Brown \cite[Theorem 8.5.7]{vzb:stacky}.  Lemma ~\ref{lem:raise-stacky-order} implies that if the main result, Theorem 
~\ref{thm:main}, holds for a curve with signature
$(g;e_1', \ldots, e_k', e_{k+1}' \ldots, e_r';\delta)$ with $e_{k+1}' = \cdots = e_r'$ satisfying certain admissibility (cf. Definition ~\ref{defn:admissible}), then Theorem ~\ref{thm:main} also holds for a curve with
signature $(g; e_1', \ldots, e_k',e_{k+1}'+2,\ldots, e_r'+2; \delta)$. 
%In order to state Lemma
%~\ref{lem:raise-stacky-order}, we will need to define a certain notion of admissibility in %Definition ~\ref{defn:admissible}, which is quite similar to \cite[Definition 8.5.1]{vzb:stacky}.

One key difference between the notion of admissibility in Definition ~\ref{defn:admissible} and that of Voight and Zureick-Brown \cite[Definition 8.5.1]{vzb:stacky} is that 
%in \cite{vzb:stacky}, Voight and Zureick-Brown may assume that $\{P_i\} \cap \Supp(K_X) = \emptyset.$ 
we cannot assume that $\{P_i\} \cap \Supp(L_X) = \emptyset$, as $H^0(X, L_X)$ may have no global sections.  We circumvent this issue by working with orders of zeros and poles relative to the divisor we are inducting from, instead of relative to the structure sheaf.
